A Triple Origin for Twin Blue Stragglers in Compact Binaries

### 2018-12-29 BSS7782.tex
edited tex file

We propose a formation mechanism for twin blue stragglers in compact
binaries that involves mass transfer from an evolved outer tertiary
companion on to the inner binary via a circumbinary disk.  We apply
our scenario to the observed double blue straggler system Binary 7782
in the old open cluster NGC 188, and show that its observed properties
are naturally reproduced within the context our proposed model.
Within the context of the hypothetical formation mechanism proposed
here, the presented work predicts the following properties for the
post-mass transfer double BS tertiary: (1) For the outer tertiary
orbit, the initial orbital period should lie between 220 days
$\lesssim$ P$_{\rm out}$ $\lesssim$ 1100 days \simon{Do you think that
  the initial binary could be as large as 1400 days?}, assuming
initial masses for the inner binary components of $m_{\rm 1} = 1.1$
M$_{\odot}$ and $m_{\rm 2} =$ 0.9 M$_{\odot}$ and an outer tertiary
mass of $m_{\rm 3} = 1.4$ M$_{\odot}$.  After Roche-lobe overflow, the
outer star turns into a white dwarf (WD). There is a correlation
bewteen the mass of this WD and the outer orbital period in the sense
that a more massive white dwarfs will be on a wider orbit.  (3) The
rotational axes of the BSs in the inner binary should be aligned with
each other and the orbital plane of the outer tertiary WD. (4) The BSs
in the inner binary should have roughly equal masses, independent of
their initial masses.  This predicts that the initially lower mass MS
star should accrete the most, and should hence be enriched more
effectively by the accreted material.  This could be observable in the
surface layers of a radiative star (i.e., He, C and O enrichment if
the donor is an RGB star, and/or s-process elements if the donor is an
AGB star). (5) Twin BSs in compact binaries formed from the proposed
mechanism should be more frequent in younger clusters \simon{WHy not
  in the field?} with ages $\lesssim$ 4-6 Gyr, since the donor will
have a radiative envelope. (6) the orbital of the binary BS will be
tight (typically $\aplt 0.3$\,au) and close to circular ($e \aplt
0.2$).

@@

Blue straggler (BS) stars are brighter and bluer than the
main-sequence turn-off (MSTO) in a cluster colour-magnitude diagram
\citep[e.g.][]{1953AJ.....58...61S,simunovic14,simunovic16}.  Two
primary channels for BS formation have been proposed; mass transfer
from an evolved donor on to a main-sequence star in a binary star
system
\citep[e.g.][]{mccrea64,1997A&A...328..143P,knigge09,mathieu09,leigh11,geller11,geller12,gosnell14,gosnell15},
and direct stellar collisions involving main-sequence stars possibly
mediated via binaries
\citep[e.g.][]{hills75,1997A&A...328..130P,shara97,leigh07,leigh11,leigh13,hypki13,2018arXiv181100058P}.
Other possible, albeit related, formation mechanisms include mergers
of close main-sequence binaries (ms, ms) \cite{2018arXiv181100058P},
and mergers of the inner binaries of hierarchical triple star systems
induced by Lidov-Kozai oscillations coupled with tidal damping
\citep[e.g.][]{perets09}.

In spite of these specific predictions for the expected properties of
BSs formed from each of the above production mechanisms, there exist
many BSs with observed properties that defy these simple scenarios.
For example, in the old open cluster M67, there lurks a candidate
triple system that is posited to host two BSs
\citep{vandenberg01,sandquist03}.  The observations suggest that the
outer tertiary is itself a BS, with a mass $\sim$ 1.7 M$_{\odot}$ and
orbiting the inner binary with a period of 1188.5 days
\citep{sandquist03}.  The inner binary has a period of only 1.068 days
\citep{vandenberg01}, and hosts a BS of mass 2.52 M$_{\odot}$.  In
order to form this system \del{after convolving its observed
  properties with its host cluster age,} we require at least five
stars in order to conserve the total system mass \citep{leigh11}.  As
explained in \citet{leigh11}, this is strongly indicative of a
dynamical origin for the system, and a single direct interaction
involving a binary and a triple that resulted in two separate
collision events is the most probable explanation for its origin
(i.e., a single interaction involving two multiples with two or more
stars is a more likely scenario to produce this system than two
back-to-back direct binary-binary interactions)
\citep{2004MNRAS.350..615G}.  \simon{Maybe add something about the
  curious binary AE Aurigae?}

Even more curious, there exists in the old open cluster NGC 188 a
double BS binary, called Binary 7782.  More generally, the BS
population in NGC 188 has a bi-modal period-eccentricity distribution.
As discussed in \citet{leigh11}, this could be hinting at a triple
origin for at least some subset of the total BS population.  As for
Binary 7782, \citet{mathieu09} observed a compact and mildly eccentric
(i.e., $e \sim 0.1$) binary star system with an orbital period of
$\sim$ 10 days hosting two roughly equal-mass blue stragglers.  During
a given binary-binary interaction, the probability that not one but
two direct (ms, ms) collisions occur is less than $10^{-2}$
\citep{leonard89,leigh11,leigh12}.  Binaries with collision products
typically have relatively long orbital periods
\cite{2011Sci...334.1380F}. Dynamically, it is difficult to form a
short-period binary composed of two collision products during a
collision interaction in a star cluster \citep{2011Sci...334.1380F}.
So, how did Binary 7782 form?

We propose a formation channel for Binary 7782, and compact double BS
binaries in general, which involves mass transfer from an outer
tertiary companion on to an inner (ms, ms) binary.  In
section~\ref{sect:dyn}, we constrain the range of initial (i.e., pre-mass
transfer) orbital parameters for a hypothetical outer tertiary
companion, using a combination of dynamical and stellar
evolution-based constraints.  In Section~\ref{sims} we present the
numerical simulations used to study the mass transfer process in our
hypothetical triple system, computed using the
\texttt{AMUSE}\cite{AMUSE} software package, in order to study the
evolution of the inner and outer orbital parameters during mass
transfer.  We summarize and discuss the implications of our results
for compact double BS binaries and, more generally, mass transfer in
stellar triples in Section~\ref{sect:discussion}.

@@

Consider a hierarchical triple system with component masses $m_{\rm
  1}$ and $m_{\rm 2}$ for the inner binary, and mass $m_{\rm 3}$ for
the outer tertiary companion.  The inner and outer binary orbital
semi-major axes are denoted a$_{\rm in}$ and a$_{\rm out}$,
respectively.  For clarity we now assume both orbits, the inner as
well as the outer to have a small eccentricity and low inclination.
These assumptions are also supported by the population of observed
low-mass triples \cite{2010yCat..73890925T,moe18}.  This initial
configuration for our assumed formation scenario for Binary 7782,
described below, is depicted in Figure~\ref{fig:fig1}.

\begin{figure}[ht!]
\plotone{fig1.eps}
\caption{Cartoon depiction of our proposed scenario for the formation
  of Binary 7782, specifically mass transfer from an evolved outer
  tertiary companion on to a compact inner binary via a circumbinary
  disk.  The outer tertiary component has mass m$_{\rm 3}$, whereas
  the inner binary components have masses m$_{\rm 1}$ and m$_{\rm 2}$.
  The inner and outer orbital separations are denoted by,
  respectively, a$_{\rm in}$ and a$_{\rm out}$.  The circularization
  radius of the accretion stream is denoted a$_{\rm c}$, as calculated
  via Equation~\ref{eqn:ac}, and marks the mean separation of the
  circumbinary disk.\
  \simon{Should we also indicate the accretion stream onto the two stars?}
\label{fig:fig1}}
\end{figure}

We consider a scenario where the outer star is filling its Roche lobe
and transfers mass to the inner binary.  The mass transfer stream
gathers at the circularization radius a$_{\rm c}$, and forms a disk
around the inner binary.  Using conservation of angular momentum, we
equate the specific angular momentum of the accreted mass at the inner
Lagrangian point of the (outer) donor star to the final specific
angular momentum of the accretion stream at the circularization radius
about the inner binary, this results in
\begin{equation}
\label{eqn:specangmom1}
v_{\rm orb,3}(a_{\rm out} - R_{\rm L}) = v_{\rm orb,c}a_{\rm c},
\end{equation}
where R$_{\rm L}$ is the radius of the Roche lobe of the outer
tertiary companion, $a_{\rm c}$ is the semi-major axis of the orbit
about the inner binary corresponding to the circularization radius and
v$_{\rm orb,c}$ is the orbital velocity at $a_{\rm c}$.  The distance
from the center of mass corresponding to the outer tertiary companion
defined by the Roche lobe is given by Equation 2 in
\citep{eggleton83}.  Combining Equation 2 in \citet{eggleton83} (with
mass ratio q $=$ m$_{\rm 3}$/(m$_{\rm 1} +$m$_{\rm 2}$)) with
Equation~\ref{eqn:specangmom1}, we can solve for the circularization
radius as a function of a$_{\rm out}$ and the assumed stellar masses:
\begin{equation}
\label{eqn:ac}
a_{\rm c} = a_{\rm out}(1 - R_{\rm L}).
\end{equation}
In order for a circumbinary disk to form around the inner binary, we
require that a$_{\rm in} <$ a$_{\rm c}$.

Figure~\ref{fig:fig2} shows the parameter space in the P$_{\rm
  out}$-P$_{\rm in}$-plane for Binary 7782.  We assume initial
component masses of $m_{\rm 1} = 1.1$ M$_{\rm \odot}$ and $m_{\rm 2}
= 0.9$ M$_{\rm \odot}$ for the inner binary components, and $m_{\rm 3}
= 1.4 $M$_{\rm \odot}$ for the outer tertiary.  We compare the
circularization radius to the semi-major axis of the inner binary, for
which we require $a_{\rm c} > a_{\rm in}$, after folding in all
constraints from the requirements for dynamical stability (listed in
the caption of Figure~\ref{fig:fig2}), an outer tertiary that is Roche
lobe-filling, and a dynamically hard outer tertiary orbit.  Note that
the range of plotted orbital periods P$_{\rm in}$ corresponding to a
contact state for the inner binary lies outside the range of plotted
values for $P_{\rm in}$ (for components with radii of roughly 1
R$_{\odot}$), since it does not contribute significantly to
constraining the outer orbital properties of a hypothesized outer WD
tertiary.  The thick horizontal solid red line shows the allowed range
of outer semi-major axes, after folding in all of the aforementioned
criteria.  As is clear, this makes a relatively narrow prediction for
the allowed ranges of outer tertiary orbits, namely 2.2 $\times$
10$^{2}$ days $\le$ P$_{\rm out}$ $\le$ 1.1 $\times$ 10$^3$ days, for
our assumed final donor mass.

\begin{figure}[ht!]
\plotone{fig2.eps}
\caption{Parameter space in the P$_{\rm out}$-P$_{\rm in}$-plane
  allowed for the hypothetical outer tertiary orbit of Binary 7782.
  The solid diagonal black lines show the period corresponding to the
  circularization radius a$_{\rm c}$ for the mass transfer stream
  coming from the outer star, for different values of the mass ratio,
  namely q $=$ 0.1, 0.7, 1, 2 and 10 \simon{before or after mass
    transfer?}.  \simon{what is the purpose of extending q to 0.1 and
    10? - NL: Fair enough.  It was arbitrary, admittedly, but supposed to take in to account the fact that the mass ratio changes during MT.  I suppose I thought it illustrated that nothing much changes upon changing q, hence the conclusions are robust over the entire extent of the MT process, during which the mass ratio of course changes.} We assume initial component masses of m$_{\rm 1}$ = 1.1
  M$_{\rm \odot}$ and m$_{\rm 2}$ = 0.9 M$_{\rm \odot}$ for the inner
  binary components, and m$_{\rm 3}$ is computed for the outer
  tertiary according to our assumed mass ratio (with our fiducial case
  corresponding to q $=$ 0.7).  We assume completely conservative mass
  transfer for this exercise, and a final mass for the outer tertiary
  of 0.6 M$_{\rm \odot}$ once it has become a WD.  The dashed diagonal
  black line shows a rough criterion for dynamical stability in the
  triple, approximately following \citet{mardling99} (i.e., a$_{\rm
    in} <$ 0.1a$_{\rm out}$ is required for long-term dynamical
  stability in equal-mass co-planar triples).  The vertical solid red
  lines show the outer orbital periods corresponding to the hard-soft
  boundary assuming central velocity dispersions of $\sigma =$ 1, 5
  and 10 km s$^{-1}$.  The vertical dashed black lines show the
  maximum outer orbital period P$_{\rm out}$ for which the outer
  tertiary companion is Roche lobe-filling, assuming a stellar radius
  of R$_{\rm 3} =$ 200 R$_{\odot}$ (CHANGE, POSSIBLY USING MESA
  CALCULATIONS?!).  The horizontal dashed red line shows the observed
  orbital period for Binary 7782, using its observed orbital period
  and our assumed final inner companion masses (i.e., m$_{\rm 1} =$
  m$_{\rm 2} =$ 1.4 M$_{\odot}$).  Finally, the thick solid horizontal
  red line shows the parameter space for P$_{\rm out}$ allowed after
  considering all of the aforementioned criteria.  \simon{I think that
    the figure needs work. I like putting the parameter space in terms
    of orbital periods, but the q=0.1 and q=10 constraints seems
    not very practical. The red labels are not readable, and I am not
    really sure how to read the figure. What about the alternative in
    the next figure, which shows tertiary mass vs. orbital period? - NL: I am fine with swapping out this figure for another, ultimately.  It still does need work.  Could you send me the files you have for radius as a function of primary mass?  I can then update the figure with that, using your assumed initial masses for the simulations.  That *would* basically finish everything I had intended with the figure, so if we still want to remove it after that, that is totally fine.}
\label{fig:fig2}}
\end{figure}

\begin{figure}[ht!]
  \includegraphics[width=\columnwidth]{fig_minimumstablesize.pdf}
\caption{Orbital separation as a function of the mass of the
  Roche-lobe filling outer star. The red curves gives the size r$_{\rm 3}$ at the
  first giant branch (thick curve) and at the AGB (thin curve). The
  blue curve gives the orbital separation a$_{\rm RLOF}$ at the moment this giant fills its
  Roche-lobe assuming a circular orbit and a total inner binary mass
  of 2.0\,\MSun. The green curves give, under these conditions, the
  maximum orbital separation for a stable inner binary or a$_{\rm stable}$.
\label{fig:tertiarymass_vs_size}}
\end{figure}

In Fig.\,\ref{fig:tertiarymass_vs_size} we present..

@@

The stellar-evolution model, its structure, temperature and
composition profiles are turned into a smoothed-particles
representation using the module {\tt StellarModelInSPH} in AMUSE (see
chapter 4 in \cite{AMUSE}).  We follow the same procedure as described
in \cite{2014MNRAS.438.1909D} for simulating the future of the triple
system \xi Tau (HD 97131) in which the outer-most star overfills it
Roche lobe to an inner binary.  After generating the hydrodynamical
representation of the donor star we replace the stellar core by a
point mass to prevent the majority of the resolution to be confined in
the star's central region.  In a following step we relax the star
using the hydrodynamics solver. This relaxation process is realized in
100 steps during which we reduce the velocity dispersion of individual
SPH particles to a glasseas structure \citep[see, for example, \S\,3.3
  on page 40 in][]{1994astro.ph.10043W}. During this procedure the
gaseous envelope of the star tends to expand by about 20\%.  To
detemrine the radius of the eventual star we calculate Lagrangian
radii and use the distance to the stellar center which contains 90\%
of it's mass. This 90\% mass-radius we consider as the stellar radius,
and match it with the Roche-lobe of the outer orbit.

With these parameters the orbital separation of the outer binary
becomes $\sim 3.86$\,au, and we adopt the outer orbit to be circular.
We choose a relative initial inclination between the inner and the
outer orbit of $i = 0^\circ$, $i = 9^\circ$, $i = 90^\circ$ and $i =
180^\circ$.  The first two choices are motivated by observed stellar
triples composed of low-mass stars which have been observed to favor
co-planar configurations \citep{moe18}. To test the effect of highly
inclined and retrogade orbits we include the $i = 90^\circ$ and $i =
180^\circ$ cases.  In fig.\,\ref{fig:topview_at_t0} we present a top
view of the triple system at the onset of Roche-lobe overflow.

@@

Roche-lobe overflow in triples is modeled using a coupled integrator
to follow the complex hydrodynamics of the mass transfer from the
Roche-lobe filling outer star to the inner binary, while keeping track
of the gravitational dynamics of the stars.  The equations of motion
of the inner binary are solved using the symplectic direct N-body
integrator \texttt{Huayno} \citep{2012NewA...17..711P}. The
hydrodynamics is performed with the smoothed-particles hydrodynamics
code \texttt{Gadget2} \citep{2000ascl.soft03001S}, using an adiabatic
equation of state.  The two inner binary stars are treated as point
masses, but we allow them to accrete mass and angular momentum from
the gas liberated by the outer star.  This is realized using
sink-particles that co-move with the mass points in the gravity
code. While the inner two stars accrete mass, they also accrete the
corresponding amount of angular momentum from the gas (see chapter 5
in \cite{AMUSE}).  The the N-body integrator accounts of this
appropriately.  For the radius of the sink particles we adopt $10
R_\odot$ for both stars.

@@

In fig.\,\ref{fig:topview_at_t1000day} we show the same triple after
1000\,days of evolution.

@@

Our choice for the initial mass of the outer tertiary is critical since it
ensures that the donor star during the mass transfer process will have
a radiative envelope \citep[e.g.][]{maeder09}.  In turn, this ensures
that the mass transfer will be maximally conservative, such that the
accretion stream will be maximally stable, accreting at a stable and
roughly constant rate \citep[e.g.][]{iben91}.

>>>>

We propose a formation mechanism for twin blue stragglers in compact
binaries that involves mass transfer from an evolved outer tertiary
companion on to the inner binary via a circumbinary disk.  We apply
our scenario to the observed double blue straggler system Binary 7782
in the old open cluster NGC 188, and show that its observed properties
are naturally reproduced within the context our proposed model.
Within the context of the hypothetical formation mechanism proposed
here, the presented work predicts the following properties for the
post-mass transfer double BS tertiary: (1) For the outer tertiary
orbit, the initial orbital period should lie between 220 days
$\lesssim$ P$_{\rm out}$ $\lesssim$ 1100 days \simon{Do you think that
  the initial binary could be as large as 1400 days?}, assuming
initial masses for the inner binary components of $m_{\rm 1} = 1.1$
M$_{\odot}$ and $m_{\rm 2} =$ 0.9 M$_{\odot}$ and an outer tertiary
mass of $m_{\rm 3} = 1.4$ M$_{\odot}$.  After Roche-lobe overflow, the
outer star turns into a white dwarf (WD). There is a correlation
bewteen the mass of this WD and the outer orbital period in the sense
that more massive white dwarfs will be on wider orbits.  (3) The
rotational axes of the BSs in the inner binary should be aligned with
each other and the orbital plane of the outer tertiary WD. (4) The BSs
in the inner binary should have roughly equal masses, independent of
their initial masses.  This predicts that the initially lower mass MS
star should accrete the most, and should hence be enriched more
effectively by the accreted material.  This could be observable in the
surface layers of a radiative star (i.e., He, C and O enrichment if
the donor is an RGB star, and/or s-process elements if the donor is an
AGB star). (5) Relative to old dense clusters with high velocity dispersions, 
twin BSs in compact binaries formed from the proposed 
mechanism should be more frequent in the Galactic field and younger 
open clusters with ages $\lesssim$ 4-6 Gyr, since then the donor will
have a radiative envelope. (6) the orbit of the binary BS will be
compact (typically $\aplt 0.3$\,au) and close to circular ($e \aplt
0.2$).

@@

Blue straggler (BS) stars are brighter and bluer than the
main-sequence turn-off (MSTO) in a cluster colour-magnitude diagram
\citep[e.g.][]{1953AJ.....58...61S,simunovic14,simunovic16}.  Two
primary channels for BS formation have been proposed: mass transfer
from an evolved donor on to a main-sequence star in a binary star
system
\citep[e.g.][]{mccrea64,1997A&A...328..143P,knigge09,mathieu09,leigh11,geller11,geller12,gosnell14,gosnell15},
and direct stellar collisions involving main-sequence stars possibly
mediated via binaries
\citep[e.g.][]{hills75,1997A&A...328..130P,shara97,leigh07,leigh11,leigh13,hypki13,2018arXiv181100058P}.
Other possible, albeit related, formation mechanism includes mergers
of close main-sequence binaries (MS-MS) \cite{2018arXiv181100058P},
and mergers of the inner binaries of hierarchical triple star systems
induced by Lidov-Kozai oscillations coupled with tidal damping
\citep[e.g.][]{perets09}.

In spite of these specific predictions for the expected properties of
BSs formed from each of the above production mechanisms, there exist
many BSs with observed properties that defy these simple scenarios.
For example, in the old open cluster M67, there lurks a candidate
triple system that is posited to host two BSs
\citep{vandenberg01,sandquist03}.  The observations suggest that the
outer tertiary is itself a BS, with a mass $\sim$ 1.7 M$_{\odot}$ and
orbiting the inner binary with a period of 1188.5 days
\citep{sandquist03}.  The inner binary has a period of only 1.068 days
\citep{vandenberg01}, and hosts a BS of mass 2.52 M$_{\odot}$.  In
order to form this system \del{after convolving its observed
  properties with its host cluster age,} we require at least five
stars in order to conserve the total system mass \citep{leigh11}.  As
explained in \citet{leigh11}, this is strongly indicative of a
dynamical origin for the system, and a single direct interaction
involving a binary and a triple that resulted in two separate
collision events is the most probable explanation for its origin
(i.e., a single interaction involving two multiples with two or more
stars is a more likely scenario to produce this system than two
back-to-back direct binary-binary interactions)
\citep{2004MNRAS.350..615G}.  \simon{Maybe add something about the
  curious binary AE Aurigae? - NL2: For sure.  I still don't know anything about it... But if I get time over the next few days I will do my research and write something.}

Even more curious, there exists in the old open cluster NGC 188 a
double BS binary, called Binary 7782.  More generally, the BS
population in NGC 188 has a bi-modal period-eccentricity distribution.
As discussed in \citet{leigh11}, this could be hinting at a triple
origin for at least some subset of the total BS population.  As for
Binary 7782, \citet{mathieu09} observed a compact and mildly eccentric
(i.e., $e \sim 0.1$) binary star system with an orbital period of
$\sim$ 10 days hosting two roughly equal-mass blue stragglers.  During
a given binary-binary interaction, the probability that not one but
two direct (MS-MS) collisions will occur is less than $10^{-2}$
\citep{leonard89,leigh11,leigh12}.  Plus, binaries with collision products
typically have relatively long orbital periods
\cite{2011Sci...334.1380F}. Dynamically, it is difficult to form a
short-period binary composed of two collision products during a
collisional interaction in a star cluster \citep{2011Sci...334.1380F}.
So, how did Binary 7782 form?

We propose a formation channel for Binary 7782, and compact double BS
binaries in general, which involves mass transfer from an outer
tertiary companion on to an inner (MS-MS) binary.  In
section~\ref{sect:dyn}, we constrain the range of initial (i.e., pre-mass
transfer) orbital parameters for a hypothetical outer tertiary
companion, using a combination of dynamical and stellar
evolution-based constraints.  In Section~\ref{sims} we present the
numerical simulations used to study the mass transfer process in our
hypothetical triple system, computed using the
\texttt{AMUSE}\cite{AMUSE} software package, in order to study the
evolution of the inner and outer orbital parameters during mass
transfer.  We summarize and discuss the implications of our results
for compact double BS binaries and, more generally, mass transfer in
stellar triples in Section~\ref{sect:discussion}.

@@

Consider a hierarchical triple system with component masses $m_{\rm
  1}$ and $m_{\rm 2}$ for the inner binary, and mass $m_{\rm 3}$ for
the outer tertiary companion.  The inner and outer binary orbital
semi-major axes are denoted a$_{\rm in}$ and a$_{\rm out}$,
respectively.  For clarity we now assume both orbits, the inner as
well as the outer, to have a small eccentricity and low inclination.
These assumptions are also supported by the population of observed
low-mass triples \cite{2010yCat..73890925T,moe18}.  This initial
configuration for our assumed formation scenario for Binary 7782,
described below, is depicted in Figure~\ref{fig:fig1}.

\begin{figure}[ht!]
\plotone{fig1.eps}
\caption{Cartoon depiction of our proposed scenario for the formation
  of Binary 7782, specifically mass transfer from an evolved outer
  tertiary companion on to a compact inner binary via a circumbinary
  disk.  The outer tertiary component has mass m$_{\rm 3}$, whereas
  the inner binary components have masses m$_{\rm 1}$ and m$_{\rm 2}$.
  The inner and outer orbital separations are denoted by,
  respectively, a$_{\rm in}$ and a$_{\rm out}$.  The circularization
  radius of the accretion stream is denoted a$_{\rm c}$, as calculated
  via Equation~\ref{eqn:ac}, and marks the mean separation of the
  circumbinary disk.\
  \simon{Should we also indicate the accretion stream onto the two stars? - NL2:  I have now updated this plot to show two circumstellar disks, and the flow on to thiose.  I can edit further, if you recommend it, but it's been tough with the down-sizing.  Maybe if I make things bigger still...  Anyways, happy to try more.}
\label{fig:fig1}}
\end{figure}

We consider a scenario where the outer star is filling its Roche lobe
and transfers mass to the inner binary.  The mass transfer stream
gathers at the circularization radius a$_{\rm c}$, and forms a circumbinary disk
around the inner binary.  Using conservation of angular momentum, we
equate the specific angular momentum of the accreted mass at the inner
Lagrangian point of the (outer) donor star to the final specific
angular momentum of the accretion stream at the circularization radius
about the inner binary, this results in
\begin{equation}
\label{eqn:specangmom1}
v_{\rm orb,3}(a_{\rm out} - R_{\rm L}) = v_{\rm orb,c}a_{\rm c},
\end{equation}
where R$_{\rm L}$ is the radius of the Roche lobe of the outer
tertiary companion, $a_{\rm c}$ is the semi-major axis of the orbit
about the inner binary corresponding to the circularization radius and
v$_{\rm orb,c}$ is the orbital velocity at $a_{\rm c}$.  The distance
from the center of mass corresponding to the outer tertiary companion
defined by the Roche lobe is given by Equation 2 in
\citep{eggleton83}.  Combining Equation 2 in \citet{eggleton83} (with
mass ratio q $=$ m$_{\rm 3}$/(m$_{\rm 1} +$m$_{\rm 2}$)) with
Equation~\ref{eqn:specangmom1}, we can solve for the circularization
radius as a function of a$_{\rm out}$ and the assumed stellar masses:
\begin{equation}
\label{eqn:ac}
a_{\rm c} = a_{\rm out}(1 - R_{\rm L}).
\end{equation}
In order for a circumbinary disk to form around the inner binary, we
require that a$_{\rm in} <$ a$_{\rm c}$.

Figure~\ref{fig:fig2} shows the parameter space in the P$_{\rm
  out}$-P$_{\rm in}$-plane for Binary 7782.  We assume initial
component masses of $m_{\rm 1} = 1.1$ M$_{\rm \odot}$ and $m_{\rm 2}
= 0.9$ M$_{\rm \odot}$ for the inner binary components, and $m_{\rm 3}
= 1.4 $M$_{\rm \odot}$ for the outer tertiary.  We compare the
circularization radius to the semi-major axis of the inner binary, for
which we require $a_{\rm c} > a_{\rm in}$, after folding in all
constraints from the requirements for dynamical stability (listed in
the caption of Figure~\ref{fig:fig2}), an outer tertiary that is Roche
lobe-filling, and a dynamically hard outer tertiary orbit.  Note that
the range of plotted orbital periods P$_{\rm in}$ corresponding to a
contact state for the inner binary lies outside the range of plotted
values for $P_{\rm in}$ (for components with radii of roughly 1
R$_{\odot}$), since it does not contribute significantly to
constraining the outer orbital properties of a hypothesized outer WD
tertiary).  The thick horizontal solid red line shows the allowed range
of outer semi-major axes, after folding in all of the aforementioned
criteria.  As is clear, this makes a relatively narrow prediction for
the allowed ranges of outer tertiary orbits, namely 2.2 $\times$
10$^{2}$ days $\le$ P$_{\rm out}$ $\le$ 1.1 $\times$ 10$^3$ days, for
our assumed final donor mass.

\begin{figure}[ht!]
\plotone{fig2.eps}
\caption{Parameter space in the P$_{\rm out}$-P$_{\rm in}$-plane
  allowed for the hypothetical outer tertiary orbit of Binary 7782.
  The solid diagonal black line shows the period corresponding to the
  circularization radius a$_{\rm c}$ for the mass transfer stream
  coming from the outer star (i.e., at the onset of mass transfer), assuming a mass ratio of q $=$ 0.7 and component masses 
  of m$_{\rm 1} =$ 0.9 M$_{\odot}$, m$_{\rm 2} =$ 1.1 M$_{\odot}$ and 
  m$_{\rm 3} =$ 1.4 M$_{\odot}$. \simon{what is the purpose of extending q to 0.1 and
    10? - NL2: I was originally thinking that q changes during MT, but I have now removed the other lines and just left the one corresponding to our test case.} We assume initial component masses of m$_{\rm 1}$ = 1.1
  M$_{\rm \odot}$ and m$_{\rm 2}$ = 0.9 M$_{\rm \odot}$ for the inner
  binary components, and m$_{\rm 3}$ is computed for the outer
  tertiary according to our assumed mass ratio (with our fiducial case
  corresponding to q $=$ 0.7).  We assume completely conservative mass
  transfer for this exercise, and a final mass for the outer tertiary
  of 0.6 M$_{\rm \odot}$ once it has become a WD.  The dashed diagonal
  black line shows a rough criterion for dynamical stability in the
  triple, approximately following \citet{mardling99} (i.e., a$_{\rm
    in} <$ 0.1a$_{\rm out}$ is required for long-term dynamical
  stability in equal-mass co-planar triples).  The vertical solid red
  lines show the outer orbital periods corresponding to the hard-soft
  boundary assuming central velocity dispersions of $\sigma =$ 1, 5
  and 10 km s$^{-1}$.  The vertical dashed black lines show the
  maximum outer orbital period P$_{\rm out}$ for which the outer
  tertiary companion is Roche lobe-filling, assuming a stellar radius
  of R$_{\rm 3} =$ 200 R$_{\odot}$.  The horizontal dashed red line shows the observed
  orbital period for Binary 7782, using its observed orbital period
  and our assumed final inner companion masses (i.e., m$_{\rm 1} =$
  m$_{\rm 2} =$ 1.4 M$_{\odot}$).  Finally, the thick solid horizontal
  red line shows the parameter space for P$_{\rm out}$ allowed after
  considering all of the aforementioned criteria.  \simon{I think that
    the figure needs work. I like putting the parameter space in terms
    of orbital periods, but the q=0.1 and q=10 constraints seems
    not very practical. The red labels are not readable, and I am not
    really sure how to read the figure. What about the alternative in
    the next figure, which shows tertiary mass vs. orbital period? - NL2: I have now edited this figure, removed the other q curves, increased font size, etc., but am okay with it if you want to remove it for something else.}
\label{fig:fig2}}
\end{figure}

%NL2: Can we also plot the radii at the terminal-age RGB and AGB, or some maximum radius during each evolutionary phase?  
%I know this will make the plot busier, but these some more relevant to me for mass transfer than the zero-age RGB and 
%AGB radii.  What do you think?
\begin{figure}[ht!]
  \includegraphics[width=\columnwidth]{fig_minimumstablesize.pdf}
\caption{Orbital separation as a function of the mass of the
  Roche-lobe filling outer star. The red curves give the stellar radius r$_{\rm 3}$ at the
  first giant branch (thick curve) and at the AGB (thin curve). The
  blue curves give the orbital separation a$_{\rm RLOF}$ at the moment this giant fills its
  Roche-lobe assuming a circular orbit and a total inner binary mass
  of 2.0\,\MSun. The green curves give, under these conditions, the
  maximum orbital separation for a stable inner binary, or a$_{\rm stable}$.
\label{fig:tertiarymass_vs_size}}
\end{figure}

%NL2: I tried to fill in the text below, but please go through and edit it as you see fit.  Also, we need to decide which figures we are going to keep, since I am guessing there is a limit for ApJL.  This of course relates to the current Figure 2, which you may feel should be removed and would be fine with me.
In Fig.\,\ref{fig:tertiarymass_vs_size} we present several characteristic sizes and orbital separations for the problem at hand.  
The red curves show the stellar radius upon reaching the first giant branch (thick curve) and the asymptotic 
giant branch (thin curves).  The blue curves show the orbital separation at the moment of Roche-lobe overflow for each of the aforementioned giants, 
assuming a circular orbit and a total inner binary mass of $0.9 + 1.1$ M$_{\odot}$ $=$ 2.0 M$_{\odot}$.  The green curves show 
for these assumptions the maximum orbital separation for a stable inner binary pair.

@@

The stellar-evolution model, including the structure, temperature and
composition profiles, are turned into a smoothed-particles
representation using the module {\tt StellarModelInSPH} in AMUSE (see
chapter 4 in \cite{AMUSE}).  We follow the same procedure as described
in \cite{2014MNRAS.438.1909D} for simulating the future of the triple
system $\chi$ Tau (HD 97131) in which the outer-most star overfills its
Roche lobe and trnsfers mass to an inner binary.  After generating the hydrodynamical
representation of the donor star we replace the stellar core by a
point mass to prevent the majority of the resolution to be confined in
the star's central regions.  In a following step we relax the star
using the hydrodynamics solver. This relaxation process is realized in
100 steps during which we reduce the velocity dispersion of individual
SPH particles to a glasseas structure \citep[see, for example, \S\,3.3
  on page 40 in][]{1994astro.ph.10043W}. During this procedure the
gaseous envelope of the star tends to expand by about 20\%.  To
detemrine the radius of the evolving star we calculate Lagrangian
radii and use the distance to the stellar center which contains 90\%
of it's mass. From this 90\% mass-radius relation we obtain the stellar radius,
and match it with the Roche-lobe of the outer orbit.

With these parameters the orbital separation of the outer binary
becomes $\sim 3.86$\,au, and we adopt the outer orbit to be circular.
We choose a relative initial inclination between the inner and the
outer orbits of $i = 0^\circ$, $i = 9^\circ$, $i = 90^\circ$ and $i =
180^\circ$.  The first two choices are motivated by observed low-mass stellar
triples composed of low-mass stars which have been observed to favor
co-planar configurations \citep{moe18}. To test the effect of highly
inclined and retrogade orbits we include the $i = 90^\circ$ and $i =
180^\circ$ cases.  In Fig.\,\ref{fig:topview_at_t0} we present a top
view of the triple system in our simulations at the onset of Roche-lobe overflow.

@@

Roche-lobe overflow in triples is modeled using a coupled integrator
to follow the complex hydrodynamics of mass transfer from the
Roche-lobe filling outer star to the inner binary, while keeping track
of the gravitational dynamics of the stars.  The equations of motion
of the inner binary are solved using the symplectic direct N-body
integrator \texttt{Huayno} \citep{2012NewA...17..711P}. The
hydrodynamics are performed with the smoothed-particles hydrodynamics
code \texttt{Gadget2} \citep{2000ascl.soft03001S}, using an adiabatic
equation of state.  The two inner binary stars are treated as point
masses, but we allow them to accrete mass and angular momentum from
the gas liberated by the outer star.  This is realized using
sink-particles that co-move with the mass points in the gravity
code. While the inner two stars accrete mass, they also accrete the
corresponding amount of angular momentum from the gas (see chapter 5
in \cite{AMUSE}).  The N-body integrator correctly accounts for this.  For the radius of the sink particles we adopt $10
R_\odot$ for both stars.

@@

In fig.\,\ref{fig:topview_at_t1000day} we show the same triple after
1000\,days of evolution.
%NL2:  We are labeling the figures differently in the text.  You label with ``fig.\,\", whereas I use ``Figure~''.  It doesn't matter to me which we use, just thought I'd highlight this so we come back to it right before submitting.

@@

We emphasize in closing that our choice for the initial mass of the outer tertiary is critical since it
ensures that the donor star during the mass transfer process will have
a radiative envelope \citep[e.g.][]{maeder09}.  In turn, this ensures
that the mass transfer will be maximally conservative, such that the
accretion stream will be maximally stable, accreting at a stable and
roughly constant rate \citep[e.g.][]{iben91}.
